### 2018-11-21 doc/Paper.tex
Update PDF to 8e3fa25d51471df5f8313ae186ee450073dac8cd

\documentclass[runningheads]{llncs}

\usepackage{amsmath}
\usepackage{calc}
\usepackage{csquotes}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{pbox}

\captionsetup{compatibility=false}

%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stmaryrd}
\DeclareFontFamily{OT1}{cmtex}{}
\DeclareFontShape{OT1}{cmtex}{m}{n}
  {<5><6><7><8>cmtex8
   <9>cmtex9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
\DeclareFontShape{OT1}{cmtex}{m}{it}
  {<-> ssub * cmtt/m/it}{}
\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
  {<5><6><7><8>cmtt8
   <9>cmbtt9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
\DeclareFontShape{OT1}{cmtex}{bx}{n}
  {<-> ssub * cmtt/bx/n}{}
\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU

\newcommand{\Sp}{\hskip.33334em\relax}
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

\EndFmtInput
\makeatother
%
%

%%format .*.       = "\mathbin{.\!\!*\!\!.}"
%%format .=.       = "\mathbin{.\!\!=\!\!.}"
%%format .<.       = "\mathbin{.\!\!<\!\!.}"
%%format sp = " "
%%format qq = "''\!"
%%format .$$ = ".\$_m"
%%format parent' = "parent_m"
%%format  <*>       = "\mathbin{\text{\small\ttfamily{<*>}}}"
%%format  <$>       = "\mathbin{\text{\small\ttfamily{<\$>}}}"
%%format t1 = "t_1"
%%format t2 = "t_2"
%%format constructor' = "constructor_m"
%%%%format Leaf = "{\sf Leaf}"
%%%%format Fork = "{\sf Fork}"
%%format Fork3 = "{\sf Fork3}"
%%%%%%format Root = "{\sf Root}"
%%format CRoot = "C_{\mathit{Root}}"
%%%%%%"{\sf C_{Root}}"
%%format CLeaf = "C_{\mathit{Leaf}}"
%%%%%"{\sf C_{Leaf}}"
%%format CFork = "C_{\mathit{Fork}}"
%%%%% "{\sf C_{Fork}}"
%%format lexeme_Leaf = "lexeme_{\sf Leaf}"
%%format MemoRoot = "{\sf Memo_{Root}}"
%%format MemoFork = "\mathit{Fork}_m"
%%%%%%"{\sf Fork}_m"
%%%%%% "{\sf Memo_{Fork}}"
%%format MemoLeaf = "\mathit{Leaf}_m"
%%%%%%"{\sf Leaf}_m"
%%%%%% "{\sf Memo_{Leaf}}"
%%format treeLookup = "lookup_{MT}"
%%%format MemoTable = "{Cache}"
%%format buildMTree = "\mathit{build}_m"
%%%%%%"build_{MT}"
%%format constructor_m = "\mathit{constructor}_m"
%%format tree_m = "\mathit{tree}_m"
%%format left_m = "\mathit{left}_m"
%%format right_m = "\mathit{right}_m"
%%format constructorM_m = "\mathit{constructorM}_m"
%%format treeM_m = "\mathit{treeM}_m"
%%format leftM_m = "\mathit{leftM}_m"
%%format rightM_m = "\mathit{rightM}_m"
%%format up_m = "\mathit{up}_m"
%%format down_m = "\mathit{down}_m"
%%format modify_m = "\mathit{modify}_m"
%%format mkAG_m = "mkAG_m"
%%format Cxt_m = "\mathit{Cxt}_m"
%%format Root_m = "Root_m"
%%format Top_m = "Top_m"
%%format L_m = "L_m"
%%format R_m = "R_m"
%%format MemoTree = "\mathit{Tree}_m"
%%format ZipperMemoTree = "\mathit{Zipper}_m"
%%format MemoAGTree = "\mathit{AGTree}_m"
%%format C_Memo_RootLet    = "C_{\mathit{Memo\_RootLet   }}"
%%format C_Memo_Let        = "C_{\mathit{Memo\_Let       }}"
%%format C_Memo_In         = "C_{\mathit{Memo\_In        }}"
%%format C_Memo_ConsAssign = "C_{\mathit{Memo\_ConsAssign}}"
%%format C_Memo_ConsLet    = "C_{\mathit{Memo\_ConsLet   }}"
%%format C_Memo_EmptyList  = "C_{\mathit{Memo\_EmptyList }}"
%%format C_Memo_Plus       = "C_{\mathit{Memo\_Plus      }}"
%%format C_Memo_Variable   = "C_{\mathit{Memo\_Variable  }}"
%%format C_Memo_Constant   = "C_{\mathit{Memo\_Constant  }}"
%%format lexemeConsAssign  = "lexeme_{\mathit{ConsAssign}}"
%%format lexemmeConsLet    = "lexeme_{\mathit{ConsLet}}"
%%format calculateMemo     = "calculate_{\mathit{Memo}}"
%%format errsAlgolMemo     = "errs_{\mathit{AlgolMemo}}"
%%format errorsMemo             = "errors_{\mathit{Memo}}"
%%format algol68                = "algol_{\mathit{68}}"
%%format ata_algol68            = "ata\_algol_{\mathit{68}}"
%%format LetSem.constructorMemo = "LetSem.constructor_{\mathit{Memo}}"
%%format lexemeString           = "lexeme_{\mathit{String}}"
%%format upGetVarValueMemo      = "up_{\mathit{GetVarValueMemo}}"
%%format lexemeInt              = "lexeme_{\mathit{Int}}"
%%format Child1              = "Child_{\mathit{1}}"
%%format Child2              = "Child_{\mathit{2}}"
%%format Child3              = "Child_{\mathit{3}}"
%%format Childi              = "Child_{\mathit{i}}"
%%format constructorMemo     = "constructor_{\mathit{Memo}}"
%%format at = "@"
%%format Algol68m               = "Algol68_{\mathit{m}}"

% Formatting
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\WithMath}[2]{{\parbox[][][b]{\widthof{#1}}{\centering$#2$}}}

% For the circular variant of repmin

% For the specification of AGs

% Tree positions

% A command for declaring todos
\newcommand{\TODO}[1]{{\color[rgb]{1,0,0}\textbf{TODO:}\textit{#1}}}

\newcommand{\ttsub}[2]{{\text{#1}_\text{#2}}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Modern type-safe embedding of attribute grammars}
\subtitle{An exercise with functional zippers in Haskell}

% TODO(twesterhout): This looks ugly... Someone, reformat it, please :)
\author{Jo{\~a}o Paulo Fernandes\inst{1}%
   \and Pedro Martins\inst{2}%
   \and Alberto Pardo\inst{3}%
   \and Jo{\~a}o Saraiva\inst{4}%
   \and Marcos Viera\inst{3}%
   \and Tom Westerhout\inst{5}%
}
\institute{%
  CISUC -- Universidade de Coimbra, Portugal \email{jpf@dei.uc.pt} \and
  University of California, Irvine, USA \email{pribeiro@uci.edu} \and
  Universidad de la  Rep\'{u}blica, Uruguay, \email{\{pardo,mviera\}@fing.edu.uy} \and
  Universidade do Minho, Portugal, \email{saraiva@di.uminho.pt} \and
  Radboud University, The Netherlands, \email{twesterhout@student.ru.nl}%
}

\date{}

\maketitle

\begin{abstract}
  % Introduction. In one sentence, what’s the topic?

  Attribute grammars are a powerful, declarative formalism to implement and
  reason about programs which, by design, are conveniently modular.

  % State the problem you tackle

  Although a full attribute grammar compiler can be tailored to specific needs,
  its implementation is highly non trivial, and its long term maintenance is a
  major endeavor.

  % Summarize (in one sentence) why nobody else has adequately answered the
  % research question yet.

  In fact, maintaining a traditional attribute grammar system is such a hard
  effort that most such systems that were proposed in the past are no longer
  active.

  % Explain, in one sentence, how you tackled the research question.

  Our approach to implement attribute grammars is to write them as first class
  citizens of a modern functional programming language.

  % In one sentence, how did you go about doing the research that follows from
  % your big idea.

  We improve a previous zipped-based attribute grammar embedding making it
  non-intrusive (i.e. no changes need to be made to the user-defined data types)
  and type-safe. On top of that, we achieve clearer syntax by using modern
  Haskell extensions.

  % As a single sentence, what’s the key impact of your research?

  We believe our embedding can be employed in practice to implement elegant,
  efficient and modular solutions to real life programming challenges.

\keywords{%
       Embedded Domain Specific Languages
  \and Zipper data structure
  \and Memoization
  \and Attribute Grammars
  \and Higher-Order Attribute Grammars
  \and Functional Programming%
}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}

  Attribute Grammars (AGs) are a declarative formalism that was proposed by
  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
  reasoning about programs in a modular and convenient way. A concrete AG
  relies on a context-free grammar to define the syntax of a language, and on
  attributes associated to